`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R

### 2021-06-10 user-guide/ms.Rmd
okay, I think I anonymized the manuscript well enough. Someone please hire me.

---
title: "`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at http://svmiller.com/peacesciencer."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
alwaays_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
  - \usepackage{colortbl}
  - \arrayrulecolor{black}
#  - \midrule\relax
---

@@

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. In a related manner, researchers may spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy, among many others---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students, in particular, face unique challenges associated with these developments. Graduate students beginning their careers in peace science to learn how scholarship informs data in peace science and  how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

`{peacesciencer}` is designed to address these problems. Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers. Researchers can use `{peacesciencer}` to create dyad-year, state-year, and even state-day data from scratch. Afterward, they can add a variety of standard information (e.g. contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. `{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts. All are done with the maximum possible transparency. The project is available for public view of Github (https://github.com/svmiller/peacesciencer). The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works. The function manuals (http://svmiller.com/peacesciencer/reference/) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the  Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

For one, researchers may invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume, for example, a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create dyad-year and state-year data from scratch. The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume, for example, a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of `{peacesciencer}`'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer [c.f. @bowersvoors2016hiyr]. This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w], scrape *any* type of data from the internet [@boehmke2016dwr], and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m]. `{peacesciencer}` embraces this. This R programming language package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

@@

Finally, the creation and presentation of data in peace science should be 100% robust and transparent, which `{peacesciencer}` takes seriously in the following ways. The website for `{peacesciencer}` has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. `{peacesciencer}` subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge).  The project's Github contains a publicly available `data-raw` directory that shows how every data set included (and processed) in `{peacesciencer}` was created. The function manuals included `{peacesciencer}` contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way. Users can also use the project's Github to point out bugs, ask for further clarification, and propose additions. `{peacesciencer}` takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

@@

```{r, echo=F, eval=T}
tribble(~data, ~description, ~type, ~func,
        "(CoW, G-W) Dyad-Year Data", "Create dyad-years from state system membership data", "", "create_dyadyears()",
        "(CoW, G-W) State-Year Data", "Create state-years from state system membership data", "", "create_stateyears()",
        "(CoW, G-W) State-Day Data", "Create state-days from state system membership data", "", "create_statedays()",
        "Archigos (v. 4.1)", "Add Archigos political leader information", "D, S", "add_archigos()",
        "ATOP Alliances (v. 5.0)", "Add ATOP alliance data", "D", "add_atop_alliance()",
        "Capital-to-Capital Distances", "Add capital-to-capital distance", "D, S", "add_capital_distance()",
        "Convert CoW codes to G-W codes", "Add matching CoW codes for G-W states by year", "D, S", "add_ccode_to_gw()",
        "Convert G-W codes to CoW codes", "Add matching G-W codes for CoW states by year", "D, S", "add_gwcode_to_cow()",
        "CoW National Capabilities (v. 5.0)", "Adds capabilities of states by year", "D, S", "add_nmc()",
        "CoW Direct Contiguity (v. 3.2)", "Adds contiguity information by year", "D, S", "add_contiguity()",
        "CoW Alliances (v. 4.1)", "Adds CoW alliance data by year", "D", "add_cow_alliance()",
        "CoW IGOs (v. 3.0)", "Adds CoW IGO data by year", "D, S", "add_igos()",
        "CoW Major Powers (v. 2016)", "Adds CoW major power by year", "D, S", "add_cow_majors()",
        "CoW MIDs (v. 5.0)", "Add CoW MID information to data", "D", "add_cow_mids()",
        "CoW Trade Data (v. 4.0)", "Add CoW trade information to data", "D, S", "add_cow_trade()",
        "CoW Wars", "Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data", "D", "add_cow_wars()",
        "CREG (v. 1.02)", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
        "Democracy", "Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
        "GML MIDs (v. 2.1.1)", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
        "Minimum Distance (v. 2016)", "Adds minimum distance data by year", "D, S", "add_minimum_distance()",
        "Peace Years", "Calculate peace years for various disputes", "D, S", "add_peace_years()",
        "Rugged/Mountainous Terrain", "Add 'ruggedness'/'mountainous' terrain data by year", "D, S", "add_rugged_terrain()",
        "(Surplus/Gross) Domestic Product", "Add Anders et al. (2020) GDP/SDP simulations by year", "D, S", "add_sdp_gdp()",
        "Strategic Rivalries", "Add Thompson and Dreyer (2012) strategic rivalry data by year", "D, S", "add_strategic_rivalries()",
        "UCDP Armed Conflicts", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
        ) -> included

@@

`{peacesciencer}`'s functions work out of the box, though it works best with one additional package.  This is `{tidyverse}`, itself a set of packages that share a common form and design [@wickhamgrolemund2017rds]. `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality. A user can open an R session and install both packages as follows.

@@

`{peacesciencer}` has a simple "pipe"-based workflow. The top of the "pipe" is the first function to create the data. These options include `create_dyadyears()`, `create_stateyears()`, and `create_statedays()`. All options include two arguments: `system` and `mry`. `system` takes one of two values---"cow" or "gw"---for what state system membership data the user wants. If `system` is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem. `mry`---short for "most recent year"---determines whether or not the function extends the membership data to the most recently concluded calendar year. The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017). `create_dyadyears()` has a third argument---`directed`---that determines whether the user wants directed dyad-year data or non-directed dyad-year data. The function defaults to `TRUE`, returning directed dyad-year data.

@@

For example, assume a researcher wanted to replicate something analogous to Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis about the causes of conflict onset. `{peacesciencer}` can do something like this in the following pipe-based workflow. First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016. This is effectively the same function used above, but with two user-specified overrides.

@@

Next, the pipe-based workflow continues with `add_gml_mids(keep = NULL)`, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.[^keepnull] `add_peace_years()` will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.[^onpy] `add_nmc()` will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. `add_democracy()` will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. `add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether ther was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, `add_sdp_gdp()` will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by @andersetal2020bbgb.

@@

`{peacesciencer}` is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with `create_stateyears(system = 'gw')`, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019, we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`. Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states. `add_ucdp_acd()` takes three arguments: `type`, `issue`, and `only_wars`. `type` is an optional argument for the type of armed conflicts for which the researcher wants information. Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate"). If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g. `type = c("intrastate", "II")`). `issue` is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include "territory", "government", and "both". If no `issue` is specified, the function returns information for all conflicts irregarding the particular issue. `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`. The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE))` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]

@@

Finally, a skeptical reader should not think that making the process as simple as possible necessarily facilitates poor decision-making by the user. In cases where it is evident what the user wants (e.g. an estimate of the level of democracy in the state-year), `{peacesciencer}` does the necessary work to provide the user that information. However, the package makes sure to leave important decision-making to the researcher. For example, `add_cow_alliance()` returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge. `add_contiguity()` returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold. The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g. the contiguity data are not ordinal and should not be treated as such), how and where data differ (e.g. how CoW state system data differ from Gleditsch-Ward data), and how researchers should consider optimally using its functionality (e.g. `add_ucdp_acd()` probably should not lump all forms of conflict together). `{peacesciencer}` does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this. Instead, `{peacesciencer}` reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.

>>>>

---
title: "`{peacesciencer}`:  A User’s  Guide  for  Quantitative  Peace Science in R"
thanks: "Replication files are available on the author's Github account (https://github.com/svmiller/peacesciencer). **Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: steven.v.miller@gmail.com."
author:
- name: Steven V. Miller
  affiliation: Clemson University
bibliography: "`r paste0(Sys.getenv('HOME'),'/Dropbox/master.bib')`" # This would be more general
# bibliography: inst/refs.bib
abstract: "This article introduces `{peacesciencer}`, an R package that contains a litany of tools for creating data of widespread interest to the peace science community. The package is cross-platform, assuming only a somewhat recent installation of the R programming language with some of the enhanced functionality of the broadly popular `{tidvyerse}` packages. Peace science researchers can use this package to greatly reduce the time needed to perfectly recreate dyad-year and state-year data from scratch and merge in common indicators included in almost every dyad-year or state-year analysis (e.g. democracy data, contiguity data). The software is freely available on CRAN and maintains an active website documenting its features at `r params$ps_website`."
keywords: "software, statistical analysis, peace science, reproducibility"
geometry: margin=1in
csl: inst/jpr.csl
mainfont: cochineal
#sansfont: Linux Biolinum O 
fontsize: 11pt
endnote: no
# pandocparas: TRUE
sansitup: FALSE
params:
  anonymous: ""
  doublespacing: ""
  removetitleabstract: ""
  ps_website: ""
  ps_manuals: ""
  ps_github: ""
anonymous: "`r params$anonymous`"
doublespacing: "`r params$doublespacing`"
removetitleabstract: "`r params$removetitleabstract`"
alwaays_allow_html: TRUE
header-includes:
  - \usepackage{longtable}
  - \usepackage{booktabs}
  - \LTcapwidth=.95\textwidth
  - \linespread{1.05}
  - \usepackage{hyperref}
  - \usepackage{setspace}
  - \let\BeginKnitrBlock\begin \let\EndKnitrBlock\end
  - \usepackage{colortbl}
  - \arrayrulecolor{black}
#  - \midrule\relax
---

@@

This data feature tackles a recurring problem for researchers in the peace science community. True research reproducibility is achieved creating original dyad-year or state-year data from scratch, though no published guide exists that informs how to do this on their own. Instead, researchers may end up reusing old code generated in past studies, leaving them to spend time and energy adjusting the sample of states, the temporal domain, and doing whatever additional troubleshooting may arise from this practice. In a related manner, researchers may spend too much time reproducing old code for standard information that goes into any dyad-year or state-year analysis---like contiguity relationships and democracy, among many others---and have to do additional troubleshooting for how these various data sources treat missing data or treat state codes in a manner inconsistent with the more accessible Correlates of War or Gleditsch-Ward state codes. This is all compounded by changes in technology that treat the creation of the data and the analysis of data as a continuous process in which the contemporary quantitative political scientist is increasingly becoming a computer programmer as well. Graduate students, in particular, face unique challenges associated with these developments. Graduate students beginning their careers in peace science must learn how scholarship informs data in peace science and how data inform scholarship at the same time they are needing to learn quantitative methods in a chosen software package.

`{peacesciencer}` is designed to address these problems. Built around the free and open source R programming language, `{peacesciencer}` contains a suite of data and functions for creating data of interest to researchers. Researchers can use `{peacesciencer}` to create dyad-year, state-year, and even state-day data from scratch. Afterward, they can add a variety of standard information (e.g. contiguity, alliances, major power status, GDP per capita estimates, capability estimates, and more) to these data with a simple command. This is a considerable time-saver since, in the absence of it, researchers would have to more meticulously code and transform the raw data to conform to the kind of data they want. `{peacesciencer}` comes with some data innovations as well, including a comprehensive data set on democracy by year, an original data set on capitals and capital transitions, and a function to create peace years between ongoing dyad-year or state-year conflicts. All are done with the maximum possible transparency. The project is available for public view of Github (https://github.com/svmiller/peacesciencer). The `data-raw` directory on the project's Github contains information and comments about how every data set was created and the `R` directory informs the user how each function works. The function manuals (`r params$ps_manual`) contain additional comments about what each function returns and, in appropriate cases, why it is doing what it is doing. Thus, `{peacesciencer}` not only assists a peace scientist with their research, but it does so in a manner that best conforms to the  Data Access and Research Transparency Initiative (DA-RT) initiative across all political science.

@@

For one, researchers invest too much time in the construction of a data set that faithfully captures the unit of analysis. Assume a researcher wants an original data set on all directed dyad-years for Correlates of War states for an analysis of inter-state conflict. How might one do that? The answer has never been immediately obvious. No published guide exists that shows a researcher how to create these data themselves from scratch, which is one reason why software bundles like EUGene and NewGene are attractive to researchers who primarily care about the substance of their research question. After all, EUGene's primary impetus was allowing researchers to create data sets for replication of previous studies, producing a host of data types (e.g. dyad-year, state-year, dispute-year) along the way that users can amend as they saw fit. `{peacesciencer}` is primarily born from this question about how to create dyad-year and state-year data from scratch. The underlying code that produces these data types is available online and are converted into simple functions for the ease of the researcher.

Second, researchers also invest too much time in retracing steps for peace science analyses for new projects. Assume  a researcher who had finished a state-year analysis on the correlates of civil conflict onset a couple years ago and wants to start a new project that analyzes the same outcome from a different angle (or perhaps using newer data). Under these conditions, a researcher will have to find where they stored that replication code and copy-paste it into a new directory for the new project. They may then have to change the name of some files, change some code to account for potentially new column names in the newer data, and troubleshoot instances where their old code does not perform as it once did. At its worst, this process may lead to some errors by the researcher. At best, this is tedium that spends the researcher's time they would rather invest in analyzing the data. The lion's share of `{peacesciencer}`'s functionality is both creating the units of analysis for the researcher and merging in different forms of data in wide use in the peace science community so that the researcher can spend less of their time on tedium.

@@

Fourth, it is increasingly the case that as the steps between creating data and analyzing data decrease in size, the lines between them blur as well. In other words, to create data is to code data and the contemporary quantitative political scientist is increasingly becoming a computer programmer [c.f. @bowersvoors2016hiyr]. This is happening concurrent to innovations in programming languages for statistical analysis, especially the R programming language that `{peacesciencer}` uses. There have been significant advances in add-on packages that allow users to do things like get World Bank data from the internet [@arelbundock2021w], scrape *any* type of data from the internet [@boehmke2016dwr], and even format results from a statistical model for presentation in a way the reduces the probability of transcription errors to almost zero [@arelbundock2021m]. `{peacesciencer}` embraces this. This R package reduces the time required to create peace science data for analysis and also informs the user about the code required to create the kind of data the user wants.

@@

Finally, the creation and presentation of data in peace science should be 100% robust and transparent, which `{peacesciencer}` takes seriously in the following ways. The website for `{peacesciencer}` has several vignettes that describe its processes in some detail. These include how it provides reasonable estimates of democracy that may not be available in the Polity data or the Varieties of Democracy data and how a researcher can whittle dyadic dispute-years into true dyad-years through reasonable case exclusions. `{peacesciencer}` subjects itself to a battery of tests before publishing updates, making sure new features do not create duplicate entries in the original data (which is the surest sign of a botched merge).  The project's Github contains a publicly available `data-raw` directory that shows how every data set included (and processed) in `{peacesciencer}` was created. The function manuals included `{peacesciencer}` contain ample documentation that clarifies what the function is doing, what it returns to the user, and why it is doing it this way. Researchers can also use the project's Github to point out bugs, ask for further clarification, and propose additions. `{peacesciencer}` takes seriously the Data Access and Research Transparency Initiative (DA-RT) initiative across all political science and endeavors for maximum transparency, leveraging open source and version control software to inform users of what data it uses and how it uses the data.

@@

```{r, echo=F, eval=T}
tribble(~data, ~description, ~type, ~func,
        "(CoW, G-W) Dyad-Year Data", "Create dyad-years from state system membership data", "", "create_dyadyears()",
        "(CoW, G-W) State-Year Data", "Create state-years from state system membership data", "", "create_stateyears()",
        "(CoW, G-W) State-Day Data", "Create state-days from state system membership data", "", "create_statedays()",
        "Archigos (v. 4.1)", "Add Archigos political leader information", "D, S", "add_archigos()",
        "ATOP Alliances (v. 5.0)", "Add ATOP alliance data", "D", "add_atop_alliance()",
        "Capital-to-Capital Distances", "Add capital-to-capital distance", "D, S", "add_capital_distance()",
        "Convert CoW codes to G-W codes", "Add matching CoW codes for G-W states by year", "D, S", "add_ccode_to_gw()",
        "Convert G-W codes to CoW codes", "Add matching G-W codes for CoW states by year", "D, S", "add_gwcode_to_cow()",
        "CoW National Capabilities (v. 5.0)", "Adds capabilities of states by year", "D, S", "add_nmc()",
        "CoW Direct Contiguity (v. 3.2)", "Adds contiguity information by year", "D, S", "add_contiguity()",
        "CoW Alliances (v. 4.1)", "Adds CoW alliance data by year", "D", "add_cow_alliance()",
        "CoW IGOs (v. 3.0)", "Adds CoW IGO data by year", "D, S", "add_igos()",
        "CoW Major Powers (v. 2016)", "Adds CoW major power by year", "D, S", "add_cow_majors()",
        "CoW MIDs (v. 5.0)", "Add CoW MID information to data", "D", "add_cow_mids()",
        "CoW Trade Data (v. 4.0)", "Add CoW trade information to data", "D, S", "add_cow_trade()",
        "CoW Wars", "Add CoW inter-state (v. 4.0) and intra-state (v. 4.1) war data", "D", "add_cow_wars()",
        "CREG (v. 1.02)", "Add ethnic/religious fractionalization/polarization by year", "D, S", "add_creg_fractionalization()",
        "Democracy", "Add Polity (v. 2017), V-Dem (v. 10), and Marquez' (2016) UDS extensions", "D,S", "add_democracy()",
        "GML MIDs (v. 2.1.1)", "Add Gibler-Miller-Little MIDs to data", "D", "add_gml_mids()",
        "Minimum Distance (v. 2016)", "Adds minimum distance data by year", "D, S", "add_minimum_distance()",
        "Peace Years", "Calculate peace years for various disputes", "D, S", "add_peace_years()",
        "Rugged/Mountainous Terrain", "Add 'ruggedness'/'mountainous' terrain data by year", "D, S", "add_rugged_terrain()",
        "(Surplus/Gross) Domestic Product", "Add Anders et al. (2020) GDP/SDP simulations by year", "D, S", "add_sdp_gdp()",
        "Strategic Rivalries", "Add Thompson and Dreyer (2012) strategic rivalry data by year", "D, S", "add_strategic_rivalries()",
        "UCDP Armed Conflicts (v. 20.1)", "Add UCDP armed conflict data by year", "S", "add_ucdp_acd()"
        ) -> included

@@

`{peacesciencer}`'s functions work out of the box, though it works best with one additional package.  This is `{tidyverse}`, itself a suit of packages that share a common form and design [@wickhamgrolemund2017rds]. `{peacesciencer}` functions make considerable use of the component packages of `{tidyverse}` but installing and loading `{tidyverse}` will allow the researcher to make quicker use of `{peacesciencer}`'s functionality. A user can open an R session and install both packages as follows.

@@

`{peacesciencer}` has a simple "pipe"-based workflow. The top of the "pipe" is the first function to create the data. These options include `create_dyadyears()`, `create_stateyears()`, and `create_statedays()`. All options include two arguments: `system` and `mry`. `system` takes one of two values---"cow" or "gw"---for what state system membership data the user wants. If `system` is not specified, the function defaults to Correlates of War system membership data given the prominence of the Correlates of War project in the conflict data ecosystem. `mry`---short for "most recent year"---determines whether or not the function extends the membership data to the most recently concluded calendar year. The function defaults to `TRUE` in the absence of a user-specified override, meaning the data returned will include all relevant state-year data from 1816 to (as of writing) 2020 even though the Correlates of War state system data were last updated at the end of 2016 (and the Gleditsch-Ward system were last updated at the end of 2017). `create_dyadyears()` has a third argument---`directed`---that determines whether the user wants directed dyad-year data or non-directed dyad-year data. The function defaults to `TRUE`, returning directed dyad-year data.

@@

For example, assume a researcher wanted to replicate something analogous to Bremer's [-@bremer1992ddc] famous "dangerous dyads" analysis about the causes of conflict onset. `{peacesciencer}` can do something like this in the following pipe-based workflow. First, `create_dyadyears(directed = FALSE, mry = FALSE)` will create a non-directed dyad-year data frame for all Correlates of War states from 1816 to 2016.

@@

Next, the pipe-based workflow continues with `add_gml_mids(keep = NULL)`, which will add information about whether there was an ongoing dispute in the dyad (and whether it was a dispute onset) from the GML dispute data.[^keepnull] `add_peace_years()` will calculate peace years between ongoing disputes in these politically relevant non-directed dyads.[^onpy] `add_nmc()` will add information from the Correlates of War National Military Capabilities data, prominently the CINC scores that conflict scholars routinely use for assessing hypotheses of power preponderance and power parity. `add_democracy()` will add information about the level of democracy for both states in the dyad using three different indicators. Researchers can select which one they prefer from these indicators. `add_cow_alliance()` will add information about whether there was an alliance in the dyad-year by way of whether there was an agreement that included a defense pledge, a neutrality pledge, a non-aggression pledge, or an entente pledge. Finally, `add_sdp_gdp()` will add information about GDP, GDP per capita, and the surplus domestic product of both states in the dyad from the simulations done by @andersetal2020bbgb.

@@

`{peacesciencer}` is also capable of creating data for replications of standard civil conflict analyses. Suppose a researcher wants to create a state-year data frame to conduct an analysis of civil conflict onset analogous to Fearon and Laitin's [-@fearonlaitin2003eicw] well-cited analysis of civil conflict onset, but using UCDP conflict data and the Gleditsch-Ward state system for creating the appropriate universe of state-years. The pipe-based workflow will start with `create_stateyears(system = 'gw')`, creating the full universe of Gleditsch-Ward state years from 1816 to 2020. Whereas the UCDP data included in `{peacesciencer}` are from 1946 to 2019 (v. 20.1), we can subset ("filter") the data to just those years with `filter(year %in% c(1946, 2019))`. Next, we can use the `add_ucdp_acd()` function to return information about ongoing UCDP conflicts and onsets for these states. `add_ucdp_acd()` takes three arguments: `type`, `issue`, and `only_wars`. `type` is an optional argument for the type of armed conflicts for which the researcher wants information. Options include "extrasystemic", "interstate", "intrastate", and "II" (short for "internationalized intrastate"). If no `type` is specified, the function returns information about ongoing disputes and onsets for all states for all tyeps of conflict. If the user wants information about multiple types of conflict--say: intra-state wars and internationalized intra-state wars---they can specify that as a character vector (e.g. `type = c("intrastate", "II")`). `issue` is another optional argument for what issue types of conflicts the user wants beyond the type of armed conflict. Options include "territory", "government", and "both". If no `issue` is specified, the function returns information for all conflicts irregarding the particular issue. `only_wars` is an argument that subsets the data to just those with the intensity levels of "war" when `only_wars = TRUE`. The argument defaults to `FALSE`, returning information about conflicts with at least 25 deaths in addition to the conflicts with more than 1,000 deaths. In this application, `add_ucdp_acd(type = "intrastate", only_wars = FALSE))` returns state-year information about ongoing intra-state conflicts over any issue and at either of UCDP's severity thresholds.[^adducdpacdalso]

@@

Finally, a skeptical reader should not think that making the process as simple as possible necessarily facilitates poor decision-making by the user. In cases where it is evident what the user wants (e.g. an estimate of the level of democracy in the state-year), `{peacesciencer}` does the necessary work to provide the user that information. However, the package makes sure to leave important decision-making to the researcher. For example, `add_cow_alliance()` returns information about various types of alliance pledges in the dyad-year---should one exist---but leaves it to the researcher to say whether they want to define the presence of an alliance to be just a defense pledge or any type of alliance pledge. `add_contiguity()` returns information about the type of contiguity relationship in the dyad-year, but leaves it to the researcher whether they want to code a contiguity variable as the presence of a mutual land border or some other type of contiguity threshold. The documentation included in this package, and on the website, is replete with caveats about the underlying data (e.g. the contiguity data are not ordinal and should not be treated as such), how and where data issues arise (e.g. how CoW state system data differ from Gleditsch-Ward data and how one is coerced into the other), and how researchers should consider optimally using its functionality (e.g. `add_ucdp_acd()` probably should not lump all forms of conflict together). `{peacesciencer}` does not endeavor to make researchers lazy or sloppy, and it does not ultimately do this. Instead, `{peacesciencer}` reduces the tedium associated with starting quantitative peace science research, achieving this in a quick, robust, and transparent way.
